---
title: "ENVS 193DS Homework 03"
author: "Lucia Murillo"
date: "2026-05-25"
format: pdf
editor: visual
---

```{r}
#| label: setup
# reading in packages 
library(tidyverse)
library(here)
library(janitor)
library(readxl)
# creating new object from kelp data  
kelp <- read_csv("data/temp-kelp.csv")
# creating new object from personal data 
personal_data <- read_csv("data/personal-data.csv") |> 
  # cleaning names of columns 
  clean_names() |> 
  # dropping N/A values 
  drop_na(date)
```

# Problem 1: Giant Kelp fronds

## (a)

One method of determining the strength of a relationship between temperature and giant kelp frond elongation is by finding the Pearson's correlation (r), which determines the direction and strength of the linear relationship between variables. The non parametric version is the Spearman rank correlation, which determines the monotonic relationship between variables.

## (b) 
```{r}
#| label: kelp-scatterplot

# base layer ggplot
# data: kelp
ggplot(data = kelp,
       mapping = aes(
         # x-axis: temperature 
       x = temp_c,
        # y-axis: kelp elongation 
       y = kelp_elong)) + 
  # first layer: scatterplot 
  # changing color to royal blue 
  geom_point(col = "royalblue") + 
  # changing x and y axis titles, adding title 
  labs(x = "Temperature (\u00B0 C)",
       y = "Kelp elongation (cm/day)",
       title = "Kelp elongation decreases as ocean temperature increases") + 
  # changing to minimal theme 
  theme_minimal()
```

